---
title: "Write_fit_XXX_functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Write_fit_XXX_functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(fatherStay)
```

This vignette details how to write `fit_*` functions so that they integrate well with the rest of the package. There are only really two important constraints on writing custom `fit_*` functions, they are that it must take in the proper input and give out the proper outputs.

# Inputs
There are three inputs for data: `Y_train`, `X_train`, and `X_nowcast`, and an input `params` which is a named list for the parameters specific to that function. For example, a function with a parameter `foo` would be:

```{r}
fit_example <- function(Y_train, X_train, X_nowcast, params = list(foo = 2)){

}
```

# Outputs
The function must be a named list with three entries:
- `model`, which has the model created
- `prediction`, which has the predictions from the model on the nowcasting data in the format of a data frame
- `fitted_values`, which is the values from fitting the model on the training data

# Example
Suppose I wish to write a function that takes two vectors $x$ and $y$ and then creates the model $y \sim foo \cdot x$, where $foo$ is a variable that I specify. The function then uses this model to make predictions and output them. I would write this function in the following way.

```{r}
fit_example <- function(Y_train, X_train, X_nowcast, params = list(foo = 2)) {
  # create the new data frame including the desired impact of foo
  newX <- data.frame(params$foo * X_train)
  data <- cbind(Y_train, newX)

  # create the actual model
  mod <- lm(Y_train ~ ., data = data)

  # create the predictions
  colnames(X_nowcast) <- colnames(newX)
  preds <- data.frame(prediction = predict(mod, data.frame(X_nowcast)))

  # create the output
  list(model = mod, prediction = preds, fitted_values = fitted.values(mod))
}
```

And it would be used automatically by the package, or you could use it manually in the following way:

```{r}
data("iris")
Y_train <- iris$Sepal.Length[1:140]
X_train <- iris[1:140, -c(1, 5)]
X_nowcast <- iris[141:150, -c(1, 5)]

fit_example(Y_train, X_train, X_nowcast, params = list(foo = 2))
```